Supplementary Information 1

### 2025-12-16 MS2_Krill.Rmd
fix abstract

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula. In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species. Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary, and arguably more realistic, representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressing them in peripheral analyses. Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use. Such integrative approaches provide a stronger foundation for CCAMLR's objectives, particularly in managing the ecologically complex krill fishery of the Western Antarctic Peninsula.

@@

The stock–recruitment (SR) relationship was modeled using a Beverton–Holt function, following @Kinzey2019. The steepness parameter was fixed at 0.85 (phase -4) to maintain consistency with previous studies, reflecting the weak relationship between spawning biomass and recruitment in krill populations, where recruitment is largely driven by environmental variability [@Dong2025; @Kinzey2019; @Perry2020]. Unfished recruitment ($R_0$) was estimated assuming a mean recruitment of 23 (log-scale), with a prior of 10. Recruitment deviations were included to capture interannual variability using the deviation vector approach *recdev = 1*, whereby recruitment is modeled as a function of spawning stock biomass plus a deviation term. Cyclic recruitment patterns were also incorporated, assuming periodicity at three and five year intervals [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014].

@@

Following @methot2020stock, natural mortality due to predation (*M2*) was incorporate through different types of observational data types, with each informing distinct aspects of predator-prey dynamics. In our krill assessment, we incorporated predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)). Predator abundance is processed as an index of predation "effort" follows the equation:

@@

where \(M2_t\) is the predation mortality in year *t*, \(I_t\) is the predator abundance index, and \( Q \) is a catchability coefficient. Additionally, length-frequency data of krill obtained from penguin diet samples from AMLR Program are used to characterize the size-selectivity of predation. The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm. assuming that the predator can capture a wider range of krill sizes. These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size. Total natural mortality (*M*) is thus partitioned as \(M = M_1 + \sum M_2 \), where \( M_1 \) is background mortality and \( \sum M_2 \) accounts for predator-specific contributions. Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality. This dual approach—using predator abundance and diet composition—supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.

@@

Along with data on predators and environmental conditions, a fundamental input to the model was derived from fishery-dependent data and scientific monitoring programs. Fishery-dependent data were obtained from the Krill Fishery Monitoring Program under the CCAMLR SISO framework, including total reported catches, standardized CPUE (adjusted by vessel, area, and depth), and krill length-frequency distributions data were standardized using a fixed bin structure of 28 length intervals, each representing the lower edge of 0.2 cm width bins from 1.2 to 6.6 cm. Length data were assigned to partition code 0 (combined discard and retained) and sex code 0 (combined sexes), with a fixed sample size of 75 individuals per haul. This ensured consistency across years and MU and fleets (fishery, survey and predator), facilitating reliable temporal and spatial comparisons of size structure, covering the period from 1997 to 2020. Fishery-independent data originated from U.S. AMLR Program acoustic surveys included standarized biomass indices conducted between 1991 and 2011. Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) was manipulate in the same structure that fishery-dependent data and handled according to stock assessment model templates requirements (Suppl. Mat. 2). All these data, together with the previously described ecosystem variables (environmental and predator-related), along with their temporal coverage and spatial scale, are summarized in Table \@ref(tab:data-sources) and illustrated in in S1, Fig. 4.

@@

kbl(data, format = "latex", 
    booktabs = TRUE, 
    caption = "Table: Data sources used in the krill population dynamics modeling") %>%
  kable_styling(latex_options = c("hold_position", "striped"), 
                font_size = 9, 
                full_width = FALSE) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "3cm")

>>>>

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is fundamental for ecosystem-based fisheries management in the Western Antarctic Peninsula, where environmental variability and predator–prey interactions are known to strongly influence population productivity. However, most stock assessment frameworks applied to krill and other exploited marine resources are based on integrated models in which ecosystem processes are not explicitly embedded within the core population dynamics. In particular, environmental variability and predation pressure are seldom incorporated directly into the state equations governing recruitment, natural mortality, and biomass dynamics in integrated assessment models. Instead, these processes are typically explored through external sensitivity analyses or post hoc diagnostics, thereby limiting their influence on estimated stock status and management advice. In this study, an integrated length-to-age assessment model is applied to Antarctic krill, allowing ecosystem drivers to be incorporated directly into the population dynamics. Within a Stock Synthesis–based framework, alternative model configurations are evaluated in which environmental covariates and predator indices are either explicitly linked to key state processes (recruitment and natural mortality) or excluded from the structural equations. This design allows a direct comparison of their effects on recruitment, total biomass, spawning stock biomass, and fishing mortality. Systematic and coherent differences are observed among model configurations. When environmental variability and predator effects are embedded within the state equations, lower estimates of recruitment, total biomass, and spawning biomass are obtained, together with altered estimates of fishing mortality, relative to models in which ecosystem drivers are ignored. In addition, poorer statistical performance is generally observed for these ecosystem-informed models, as indicated by higher RMSE values, less stable Monod-type parameter estimates, and more structured residual patterns. This deterioration in fit is interpreted as the result of the increased informational and structural cost incurred by the model when additional ecological processes are imposed, rather than as evidence against the relevance of the ecological mechanisms themselves. Overall, these results indicate that stock assessments excluding ecosystem processes may overestimate population productivity and stock status by prioritizing statistical fit over ecological realism. Despite the challenges associated with integrating ecosystem information into fully integrated assessment models, it is demonstrated that environmental and predator effects can be consistently incorporated within the core population dynamics of krill assessments. The direct embedding of these drivers in the state equations provides a more precautionary and ecologically realistic representation of krill population dynamics and strengthens the scientific basis for management advice consistent with CCAMLR’s objectives for conservation and sustainable use in the Western Antarctic Peninsula.

@@

The stock–recruitment (SR) relationship was modeled using a Beverton–Holt function, following @Kinzey2019. The steepness parameter was fixed at 0.85 (phase -4) to maintain consistency with previous studies, reflecting the weak relationship between spawning biomass and recruitment in krill populations, where recruitment is largely driven by environmental variability [@Dong2025; @Kinzey2019; @Perry2020]. Unfished recruitment ($R_0$) was estimated assuming a mean recruitment of 23 mm. (log-scale), with a prior of 10. Recruitment deviations were included to capture interannual variability using the deviation vector approach *recdev = 1*, whereby recruitment is modeled as a function of spawning stock biomass plus a deviation term. Cyclic recruitment patterns were also incorporated, assuming periodicity at three and five year intervals [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014].

@@

Following @methot2020stock, natural mortality due to predation (*M2*) was incorporate through different types of observational data types, with each informing distinct aspects of predator-prey dynamics. In our krill assessment, we incorporated annual predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)). Predator abundance is processed as an index of predation "effort" follows the equation:

@@

where \(M2_t\) is the predation mortality in year *t*, \(I_t\) is the predator abundance index, and \( Q \) is a catchability coefficient. Additionally, length-frequency data of krill obtained from penguin diet samples from AMLR Program are used to characterize the size-selectivity of predation. The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.5 cm and a corresponding selection range width of 3.5 cm assuming that the predator can capture a wider range of krill sizes. These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size. Total natural mortality (*M*) is thus partitioned as \(M = M_1 + \sum M_2 \), where \( M_1 \) is background mortality and \( \sum M_2 \) accounts for predator-specific contributions. Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality. This dual approach (using predator abundance and diet composition) supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.

@@

Along with data on predators and environmental conditions, a fundamental input to the model was derived from fishery-dependent data and scientific monitoring programs. Fishery-dependent data were obtained from the CCAMLR Krill Fishery Monitoring Program and SISO Scheme, including total catches in weight, effort, (agruped by vessel, area, and depth), and krill length-frequency distributions data were standardized using a fixed bin structure of 28 length intervals, each representing the lower edge of 0.2 cm width bins from 1.2 to 6.6 cm. Length data were assigned to partition code 0 (combined discard and retained) and sex code 0 (combined sexes), with a fixed sample size of 75 individuals per haul. This ensured consistency across years and MU and fleets (fishery, survey and predator), facilitating reliable temporal and spatial comparisons of size structure, covering the period from 1997 to 2020. Fishery-independent data originated from U.S. AMLR Program acoustic surveys included standarized biomass indices conducted between 1991 and 2011. Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) was manipulate in the same structure that fishery-dependent data and handled according to stock assessment model templates requirements (Suppl. Mat. 2). All these data, together with the previously described ecosystem variables (environmental and predator-related), along with their temporal coverage and spatial scale, are summarized in Table \@ref(tab:data-sources) and illustrated in in S1, Fig. 4.

@@

kbl(data, format = "latex", 
    booktabs = TRUE, 
    caption = "Data sources used in the krill population dynamics modeling") %>%
  kable_styling(latex_options = c("hold_position", "striped"), 
                font_size = 9, 
                full_width = FALSE) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "3cm")

### 2025-12-16 index.Rmd
fix abstract

circleplot_res
```

@@

```{r resldens, fig.width=10, fig.height=6, fig.cap="Density of residual distribution by scenarios in krill model"}
ggplot(resid_all, aes(x = Obs - Exp, color = model)) +
  geom_density(alpha = 0.5) +
  facet_wrap(Fleet_name~type, scales = "free") +
  labs(title = "Residual Distribution", 
       x = "Residual (Obs - Exp)", 
       y = "Density") +
  scale_color_manual(name = "",
                     values = c("s1.1" = "#ca0020",  
                                "s1.2" = "#f4a582",  
                                "s1.3" = "#bababa",  
                                "s1.4" = "#404040")) +
  theme_minimal()
```
### Residual consistency

>>>>

circleplot_res

# 1. Calcular pendiente y p-value por Fleet y Modelo
lm_stats <- df_all %>%
  group_by(Model, Fleet) %>%
  do({
    mod <- lm(Pearson ~ Yr, data = .)
    out <- tidy(mod)
    slope <- out$estimate[out$term == "Yr"]
    pval <- out$p.value[out$term == "Yr"]
    data.frame(slope = round(slope, 3),
               pval = round(pval, 3))
  }) %>%
  ungroup()

# 2. Plot con nubes de puntos y línea de tendencia
ggplot(df_all, aes(x = Yr, y = Pearson)) +
  geom_point(alpha = 0.5, size = 0.7, color = "grey30") +
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 0.6) +
  facet_grid(Model ~ Fleet) +
  theme_bw(base_size = 9) +
  labs(x = "Year", y = "Pearson residuals") +
  theme(strip.text = element_text(size = 7)) +
  # 3. Añadir pendiente y p-value dentro de cada panel
  geom_text(data = lm_stats,
            aes(x = min(df_all$Yr) + 2,
                y = max(df_all$Pearson, na.rm = TRUE) - 0.3,
                label = paste0("slope=", slope, "\np=", pval)),
            inherit.aes = FALSE,
            size = 2.5, hjust = 0)

@@

```{r resldens, fig.width=10, fig.height=6, fig.cap="Density of residual distribution by scenarios in krill model"}
ggplot(resid_all, aes(x = Obs - Exp, color = model)) +
  geom_density(alpha = 0.5) +
  facet_wrap(Fleet_name~type, scales = "free") +
  labs(title = "Residual Distribution", 
       x = "Residual (Obs - Exp)", 
       y = "Density") +
  scale_color_manual(name = "",
                     values = c("s1.1" = "#ca0020",  
                                "s1.2" = "#f4a582",  
                                "s1.3" = "#bababa",  
                                "s1.4" = "#404040")) +
  theme_minimal()
```

```{r}
# Lista de nombres de modelos
model_names <- paste0("base.model1.", 1:4)

# Bucle para calcular Pearson residual
for (mn in model_names) {
  # Obtener el objeto del modelo
  mod <- get(mn)

  # Calcular residual de Pearson
  mod$cpue$PearsonRes <- with(mod$cpue, (Obs - Exp) / SE_input)

  # Guardar de vuelta en el entorno global
  assign(mn, mod)
}

# Revisar el resultado, ejemplo para el primero
head(base.model1.1$cpue[, c("Obs", "Exp", "SE_input", "PearsonRes")])

library(ggplot2)
library(dplyr)

# Lista de modelos
model_names <- paste0("base.model1.", 1:4)

# Crear un data.frame combinado con todos los modelos
cpue_all <- bind_rows(
  lapply(model_names, function(mn) {
    mod <- get(mn)
    df <- mod$cpue
    df$Model <- mn  # añadir columna con el nombre del modelo
    return(df)
  })
)

# Plot de residuos de Pearson por año, por flota, facetado por modelo
ggplot(cpue_all, aes(x = Yr, y = PearsonRes)) +
  geom_point(alpha = 0.6) +
  facet_grid(Model ~ Fleet_name) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(
    title = "Residuos de Pearson por modelo y flota",
    x = "Año",
    y = "Residual de Pearson"
  )+
  ylim( -10,10)
```

### Residual consistency
